---
title: "simPA_paper_generic"
output: html_document
date: "2024-07-22"
---

# Setup
## Libraries
```{r libraries}
library(dplyr)
library(tibble)
library(ggplot2)
library(stringr)
library(ggraph)
library(ggrepel)
library(ggnewscale)
library(igraph)
library(pals)
library(cowplot)
library(tidyr)
library(ggpubr)
library(forcats)
library(jsonlite)
library(ggh4x)
library(patchwork)
# PCA
library(FactoMineR)
library(factoextra)
library(magick)
```

## Set variables
Use the model name that is in the exported files.
Make a folder and add the path here, with a slash at the end.
```{r}
# model_name = "Recon2.2"
model_name = "Human1"
input.path = paste0("../../data/",model_name,"/r_input/")
# Set to T if using the same categories as in the pathways_db file, aka Human1 categories (used for the colour palette):
human.pathway.categories = T 

if (model_name == "Human1"){
  exch.metab.pattern = "e"
}else if (model_name == "Recon2.2"){
  exch.metab.pattern = "_e"
}
# Change these to suit your tested methods
methods.indexes = seq(1:3)
names(methods.indexes) = c("ORA", "GSEA abs", "GSEA raw")
```


## Read in data
```{r data}
# Data exported from python:
# List of all metabolites and their respective pathway
pathways = read.table(paste0(input.path,"metabolite_pathways.tsv"), sep="\t", header=T) %>% unique()
# Pathway dictionary of each pathway Human1 ID and name
pathway_dict = read.table(paste0(input.path,"pathway_dict.tsv"), sep = "\t", header=T)

# Pathway states (unable to KO, blocked, unable to ORA)
pathway.states = read.table(paste0(input.path,"pathway_states.tsv"), sep="\t", header=T)

metabolites = read.table(paste0(input.path,"metab_dict.tsv"), sep="\t", header=T,quote = "") 

# List of pathways which were unable to run because of KO infeasibility since we are using a default WT
# infeasible_pathways_ko = read.table(paste0(input.path,"infeasible_pathways.txt"), sep="\t") %>% pull(V1)
infeasible_pathways_ko = pathway.states %>% filter(Type == "infeasible") %>% pull(Pathway)

# Input data from ORA and GSEA, with a default WT for ALL pathways
# These are the output filenames from python
filenames = c(paste0(model_name,"_",list("ORA_pval_005_metrics_external", "GSEA_absolute_metric_external","GSEA_raw_metric_external"),"_with_WT_default_all"))


# Categorised pathways: this file needs to be created by hand...
pathway_categories = read.table(paste0(input.path,"pathway_db.tsv"), sep="\t", header = T)

# List of Miscellaneous pathways which are removed from the analysis
# Either a file you wrote yourself (i), or extracted from the pathway categories (ii).
# (i)
#misc_blocked_pathways = read.table(paste0(input.path,"miscellaneous_pathways.txt"), sep="\t") %>% pull(V1)
# (ii)
misc_pathways = pathway_categories %>% filter(ManualCategoryLvl1 == "Miscellaneous") %>% pull(Pathway)
blocked_pathways = pathway.states %>% filter(Type == "blocked") %>% pull(Pathway)
misc_blocked_pathways = c(misc_pathways, blocked_pathways) %>% unique()

network_stats = read.table(paste0(input.path,model_name,"_no_blocked_cytoscape_pathway_graph_statistics.csv"), sep = ",", header = T)
GSEA.abs.results = read.table(paste0(input.path,model_name,"_GSEA_absolute_metric_external_with_WT_default_all_results.tsv"), sep="\t", header=T)
z.scores = read.table(paste0("../../data/large/",model_name,"_zscores.tsv"), sep="\t", header=T) %>%
  filter(Metab %in% metabolites$ID)

side_compounds = read.table(paste0(input.path,"side_compounds.txt"), header = T) %>% 
  rename("Identifier"=1) %>% 
  mutate(Identifier = stringr::str_sub(Identifier, end=-str_length(exch.metab.pattern)-1)) %>% unique() %>% 
  mutate(Identifier = ifelse(startsWith(Identifier, "M_"), str_sub(Identifier, start=3), Identifier))
```

## Colour scheme
```{r}
basic_colour = "#56B4E9" # Used for plots where there are no other colours

# Read pathway category colours if possible:
if (human.pathway.categories){
  pathway_cat_cols = read.table(paste0("../../data/Human1/r_input/Human1_pathway_category_colours.tsv"), sep="\t", header=T, comment.char = "")
}else{
  # Otherwise generate a new palette:
  max_n = length(unique(pathway_categories$ManualCategoryLvl1))+4
  cols = glasbey()[5:max_n]
  pathway_cat_cols = pathway_categories %>% select(ManualCategoryLvl1) %>%
    unique() %>%
    mutate("col" = cols)
}

types_names = c("FN", "TP", "Infeasible", "Feasible", "Not run", 
                "Insufficient significant metabolites", "Unable to produce biomass when KO'd", "Blocked")


colours = c("#FF0000", "#3660d6","#c40d00", "#1a57ba", "#5c5c5c", "#fc86a6", "lightgrey","#808080")
dark_colours = c("#a80000", "#2d4ba1","#800800", "#203b8a", "#5c5c5c", "#e04f76","#5c5c5c","#808080")

# Base colours
pathway_cat_cols_dict = append(colours , paste0(pathway_cat_cols$col,"B3"))
names(pathway_cat_cols_dict) = append(types_names, pathway_cat_cols$ManualCategoryLvl1)

# Base colours with alpha (necessary for lighter colours in circle plots)
pathway_cat_cols_dict_light = append(colours, paste0(pathway_cat_cols$col,"66"))
names(pathway_cat_cols_dict_light) = append(types_names, pathway_cat_cols$ManualCategoryLvl1)

# Darker colours (necessary for darker outlines in circle plots)
pathway_cat_cols_line_dict = append(dark_colours , paste0(pathway_cat_cols$col, "FF"))
names(pathway_cat_cols_line_dict) = append(types_names,pathway_cat_cols$ManualCategoryLvl1)

# Add "Cat_" to the beginning of category names to avoid duplicate names with certain pathways
cat_pathway_cat_cols_dict_light = pathway_cat_cols_dict_light
names(cat_pathway_cat_cols_dict_light) = append(types_names, paste0("Cat_", pathway_cat_cols$ManualCategoryLvl1))
cat_pathway_cat_cols_line_dict = pathway_cat_cols_line_dict
names(cat_pathway_cat_cols_line_dict) = append(types_names, paste0("Cat_",pathway_cat_cols$ManualCategoryLvl1))
```


```{r}
# Read in metrics files
method_names = filenames %>% unlist()
methods_to_test = lapply(filenames, function(x){
  read.table(paste0(input.path,x,".csv"), sep = ",", header = T)
})
names(methods_to_test) = method_names

# Compile all files into one list object while converting data into a usable format
all_methods_TP_FN = lapply(methods_to_test, function(x){
  x %>%
  mutate(Type = ifelse(TPR==1,"TP", ifelse(FNR==1,"FN",""))) %>%
  filter(!Subsystem.name %in% misc_blocked_pathways) %>%
  left_join(pathway_categories, by = c("Subsystem.name" = "Pathway")) %>%
  select(Subsystem.name, ManualCategoryLvl1, Type)
})
names(all_methods_TP_FN) = filenames

pathways_to_exclude = pathway_categories$Pathway[(!pathway_categories$Pathway %in% all_methods_TP_FN[[methods.indexes[["GSEA abs"]]]]$Subsystem.name) & (!pathway_categories$Pathway %in% infeasible_pathways_ko)]

all_methods_TP_FN_cat = lapply(methods_to_test, function(x){
  not.run = methods_to_test[[methods.indexes["GSEA abs"]]]$Subsystem.name[!(methods_to_test[[methods.indexes["GSEA abs"]]]$Subsystem.name %in% x$Subsystem.name)] %>% as.data.frame() %>% 
    rename("Subsystem.name" =1) %>% 
    mutate(Type = "Insufficient significant metabolites", FPR=NA, TNR=NA) %>% 
    filter(!Subsystem.name %in% x$Subsystem.name)
  
  blocked = blocked_pathways %>% as.data.frame() %>% 
    rename("Subsystem.name" =1) %>% 
    mutate(Type = "Blocked", FPR=NA, TNR=NA)
  
  if ("TNR2" %in% colnames(x)){
    x = x %>% select(-TNR) %>% 
      rename(TNR = TNR2)
  }
  
  x %>%
    mutate(Type = ifelse(TPR==1,"TP", ifelse(FNR==1,"FN", ""))) %>%
    filter(!Subsystem.name %in% misc_blocked_pathways) %>%
    select(Subsystem.name, Type, FPR, TNR) %>% 
    rbind(infeasible_pathways_ko %>% as.data.frame() %>% rename("Subsystem.name" =1) %>% 
            mutate(Type = "Unable to produce biomass when KO'd", FPR=NA, TNR=NA)) %>% 
    rbind(not.run) %>% 
    rbind(blocked) %>% 
    filter(!Subsystem.name %in% misc_blocked_pathways) %>%
    left_join(pathway_categories %>% select(Pathway, ManualCategoryLvl1), by = c("Subsystem.name" = "Pathway")) %>%
    mutate(ManualCategoryLvl1 = paste0("Cat_", ManualCategoryLvl1))
})

# Prepare feasibility categories
all_pathway_cat_feas_infeas = pathway_categories %>% select(Pathway, ManualCategoryLvl1)  %>%
  mutate(Type = ifelse(Pathway %in% infeasible_pathways_ko, "Infeasible", 
                       ifelse(Pathway %in% misc_blocked_pathways, "Not run", "Feasible"))) %>%
  mutate(ManualCategoryLvl1 = paste0("Cat_",ManualCategoryLvl1)) %>% 
  filter(!Pathway %in% misc_blocked_pathways)

all_pathway_feas_infeas = pathway_categories %>% select(Pathway, ManualCategoryLvl1)  %>%
  mutate(Type = ifelse(Pathway %in% infeasible_pathways_ko, "Infeasible", 
                       ifelse(Pathway %in% misc_pathways, "Not run", 
                              ifelse(Pathway %in% blocked_pathways, "Blocked", "Feasible")))) %>% 
  filter(Type != "Not run")%>% 
  filter(!Pathway %in% misc_blocked_pathways)

label_dict_no_misc = c("Feasible", "Infeasible")
names(label_dict_no_misc) = c("Functional", "Infeasible")

label_dict_no_misc_swap = c("Functional", "Infeasible", "All pathways", "Blocked pathways")
names(label_dict_no_misc_swap) = c("Feasible", "Infeasible", "All","Blocked")

# Read in raw ORA and GSEA results
raw_method_results = lapply(paste0(filenames, "_results"), function(x){
    read.table(paste0(input.path,x,".tsv"), sep="\t", header=T)
})
names(raw_method_results) = method_names
method_names_clean = setNames(c("ORA", "GSEA abs","GSEA raw"), method_names)

all_methods_TP_FN_long = lapply(names(all_methods_TP_FN), function(x){
  all_methods_TP_FN[[x]] %>% select(-ManualCategoryLvl1) %>% 
    mutate(Method = x)
  })
all_methods_TP_FN_long_df = all_methods_TP_FN_long %>% bind_rows() %>% rename("Pathway" = Subsystem.name)
```

# Pathway stats
## Pathway sizes
These pathway sizes are not the same as in the default model: they do not count metabolites removed by deleting blocked reactions.
```{r}
pathway_sizes = pathways %>% group_by(subsystem) %>% summarise(size = n()) %>% rename("Pathway" = "subsystem") %>% 
  left_join(pathway_categories %>% select(Pathway, ManualCategoryLvl1), by = c("Pathway"))
pathway_sizes_inf_fea = all_pathway_feas_infeas %>% left_join(pathway_sizes, by = c("Pathway", "ManualCategoryLvl1")) %>%  rbind(pathway_sizes %>% mutate(Type = "All"))

# Colour dict
pathway_type_dict = append(pathway_cat_cols_dict[c("Infeasible", "Feasible")], basic_colour)
names(pathway_type_dict) = c("Infeasible", "Feasible", "All")

pathway_sizes_inf_fea_no_misc = pathway_sizes_inf_fea %>% 
  filter(ManualCategoryLvl1 != "Miscellaneous") 

binwidth=10
max_counts = max(hist(pathway_sizes_inf_fea_no_misc %>% filter(Type=="All") %>% pull(size), plot=F, 
                    breaks=seq(0, max(pathway_sizes_inf_fea_no_misc %>% filter(Type=="All") %>% pull(size))+21,by=binwidth))$counts)

pathway_sizes_plot =
  pathway_sizes_inf_fea_no_misc %>% 
  ggplot()+
  geom_histogram(aes(x=size, fill=Type), binwidth = binwidth)+
  facet_wrap(~Type, nrow = 3, labeller = as_labeller(label_dict_no_misc_swap))+
  theme_minimal()+
  ylab("N. of pathways")+
  xlab("Pathway size")+
  theme(strip.background = element_blank(), legend.position = "None", axis.title = element_text(size=9))+
  scale_fill_manual(values =pathway_type_dict[pathway_sizes_inf_fea$Type])+
  geom_text(data = pathway_sizes_inf_fea %>% 
  filter(ManualCategoryLvl1 != "Miscellaneous") %>% 
              group_by(Type) %>%
              slice(1) %>%ungroup() %>% 
              mutate(TypeNum = row_number(), label = LETTERS[as.numeric(TypeNum)]),
            aes(label = label, x = -45, y = max_counts*1.15), color = "black", fontface = "bold", size=5)+
    coord_cartesian(clip="off", ylim = c(0,max_counts), xlim= c(0, max(pathway_sizes_inf_fea_no_misc$size)))
pathway_sizes_plot
```

## Pathway feasible/infeasible counts per category
```{r}
category_sizes = pathway_categories %>% select(Pathway, ManualCategoryLvl1) %>% group_by(ManualCategoryLvl1) %>% summarise(n = n()) %>% arrange(n)

all_pathway_feas_infeas$ManualCategoryLvl1 = factor(all_pathway_feas_infeas$ManualCategoryLvl1, levels = category_sizes$ManualCategoryLvl1)
all_pathway_feas_infeas_blocked = all_pathway_feas_infeas %>% rbind(pathway_categories %>% filter(Pathway %in% blocked_pathways) %>% select(Pathway, ManualCategoryLvl1) %>% mutate("Type" = "Blocked"))

pathway_stats_feas_plot =
  all_pathway_feas_infeas_blocked %>%
    mutate(Type = fct_relevel(Type, c("Blocked","Feasible", "Infeasible"))) %>% 
  ggplot()+
  geom_bar(aes(y = ManualCategoryLvl1, fill = Type), position="stack")+
  theme_minimal()+
  scale_fill_manual(values=pathway_cat_cols_dict[all_pathway_feas_infeas_blocked$Type], 
                    labels = label_dict_no_misc_swap[all_pathway_feas_infeas_blocked$Type], name = "")+
  theme(axis.title = element_text(size=9),
        axis.text.y = element_text(size=9, margin = margin(r=0)),
        legend.position = "bottom",
        # legend.box="vertical", 
        # legend.position.inside = c(-20,1),
        legend.location = "plot",
     legend.justification='left',
     legend.box = "horizontal",
     # legend.margin = margin(0),
     # legend.key = element_blank(),
     legend.direction = "horizontal",
     plot.margin = margin(0.5,0.5,0.2,0.5,"cm"))+
  # guides(fill = guide_legend(nrow = 2))+
  xlab("N. of pathways")+
  ylab("")+ 
  scale_x_continuous(expand = c(0, 0))+
  scale_y_discrete(labels=label_wrap_gen(25))

pathway_stats_feas_plot
#ggsave("../../plots/stacked_barplot_r2.png", width = 8, height = 5, dpi = 300, bg = "white")
```

## Fig1
```{r}
plot_grid(
  plot_grid(NULL,pathway_sizes_plot, NULL, nrow=1, rel_widths = c(0,1,0)), 
  pathway_stats_feas_plot, labels=c("","D"), rel_widths = c(0.4,0.6), nrow=1)
ggsave(paste0("../../plots/",model_name,"/Fig1.png"),bg="white", dpi=300, width = 11, height = 5.5)
```

# Fig2
```{r}
categories_to_pathways = lapply(all_methods_TP_FN_cat, function(x){
  x %>%
  rename("from" = ManualCategoryLvl1, "to" = Subsystem.name) %>%
  relocate(from)
})

g_all = lapply(categories_to_pathways, function(x){
  g = graph_from_data_frame(x)
  g_temp = create_layout(g, 'circlepack')
  list(g, g_temp)
})

pathways_to_TP_FN = lapply(seq_along(all_methods_TP_FN_cat), function(i){
  all_methods_TP_FN_cat[[i]] %>%
  select(-ManualCategoryLvl1) %>%
  rename("Pathway" = Subsystem.name) %>%
  rbind(all_methods_TP_FN_cat[[i]] %>% select(ManualCategoryLvl1) %>% 
  unique() %>% mutate("Type" = "NA", "FPR" = "NA", "TNR" = "NA") %>% rename("Pathway" = ManualCategoryLvl1)) %>%
  arrange(match(Pathway, g_all[[i]]$g_temp$name))
})

pathway_counts = lapply(pathways_to_TP_FN, function(x){
  pathways %>% group_by(subsystem) %>%
  count() %>% rename("Pathway" = subsystem) %>%
  mutate("Name" = "subsystem") %>%
  rbind(pathway_categories%>% group_by(ManualCategoryLvl1) %>% select(Pathway, ManualCategoryLvl1) %>%
  count() %>% rename("Pathway" = ManualCategoryLvl1) %>% mutate("Name" = "category", Pathway=paste0("Cat_", Pathway))) %>%
  mutate(Name = ifelse(Name=="category", Pathway, NA)) %>%
  left_join(x %>% mutate(Name = ifelse(is.na(Type), "category", NA)), by = c("Pathway", "Name"))
})
labelled_categories = lapply(all_methods_TP_FN_cat, function(x){
  x %>% group_by(ManualCategoryLvl1) %>% 
    # count() %>%
    mutate(n = n()) %>% 
  filter(ManualCategoryLvl1 %in% paste0("Cat_",pathway_cat_cols$ManualCategoryLvl1) & (Type=="FN" |n ==1) ) %>%
  pull(ManualCategoryLvl1)
})
pathway_counts_run = lapply(seq_along(pathways_to_TP_FN), function(i){
  pathway_counts[[i]] %>% filter(Pathway %in% pathways_to_TP_FN[[i]]$Pathway) %>%
  rename("name" = Pathway) %>%
  mutate(Type = ifelse(is.na(Type), str_remove(name, "Cat_"), Type), label = ifelse(name %in% labelled_categories[[i]], str_remove(Name, "Cat_"), NA),
         linetype=ifelse(Type=="Unable to produce biomass when KO'd", 2, 
                         ifelse(Type=="Unable to produce biomass in WT",2,1)))
})

linetype_dict = ifelse(unique(pathway_counts_run[[1]]$Type) %in% 
                         c("Unable to produce biomass when KO'd", "Unable to produce biomass in WT", "Insufficient significant metabolites"), 3, 1)
names(linetype_dict) = unique(pathway_counts_run[[1]]$Type)

g = lapply(seq_along(categories_to_pathways), function(i){
  graph_from_data_frame(categories_to_pathways[[i]], vertices=pathway_counts_run[[i]])
})
layout = lapply(g, function(x){
  create_layout(x, 'circlepack', weight = n) %>% 
    mutate(r = ifelse(is.na(Name), r-0.2, r))
})
```

## GSEA raw
Run GSEA plots first to get the coordinates for all other circle plots
```{r}
base_layout = methods.indexes["GSEA raw"]
if (model_name == "Human1" | model_name == "Recon2.2"){
  layout_coord = read.table(paste0(input.path,"circlepack_layout.tsv"), sep="\t", header=T,quote = "")
  i = methods.indexes["GSEA raw"]
  layout[[i]]=layout[[i]] %>% select(-x,-y,-r) %>% left_join(layout_coord %>% select(name,x,y,r), by="name")
}else{
  i = methods.indexes["GSEA raw"]
}
```


```{r}
main_plot =
    ggraph(layout[[i]]%>% arrange(Type))+
    geom_node_circle(aes(fill=Type, color=Type, group=leaf, linetype=as.factor(Type)))+
    theme_void()+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type])+
    scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type])+
    coord_equal(clip='off')+
    theme(legend.position = "none")+
    ggraph::geom_node_label(aes(label=str_wrap(label,23)),alpha=0.6,lineheight = 0.7, size=3)

TP_FN_legend =
    layout[[i]] %>% filter(Type %in% c("TP", "FN", "Unable to produce biomass when KO'd")) %>%
    mutate(Type = fct_relevel(Type, c("TP", "FN","Unable to produce biomass when KO'd"))) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type, linetype=as.factor(Type)))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white", name="Predicted",
                      labels=c("True Positive", "False Negative", 
                               "Infeasible"))+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Predicted",
                      labels=c("True Positive", "False Negative", 
                               "Infeasible"))+
  scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type], 
                        labels=c("True Positive", "False Negative", 
                               "Infeasible"), name="Predicted")+
    theme(legend.margin = margin(c(-0.5,0,0,0)))


cat_legend = layout[[i]] %>% filter(!Type %in% c("TP", "FN", "Unable to produce biomass when KO'd")) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white", name="Pathway category")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Pathway category")+
    theme(legend.margin = margin(c(-0.5,0,0,0)))
  
p2 = plot_grid(
    main_plot,
    plot_grid(get_legend(TP_FN_legend),
    get_legend(cat_legend),nrow = 2, align = "hv", rel_heights = c(0.3,0.7))
    ,nrow = 1, rel_widths = c(0.6,0.4))
p2
  
GSEA_raw_no_leg = main_plot
```

## GSEA abs
Match circle coords from GSEA raw
```{r}
i=methods.indexes["GSEA abs"]
GSEA_abs_layout = layout[[i]] %>% select(-x,-y, -r) %>% left_join(layout[[base_layout]] %>% select(x, y,r, name), by="name")
# Layout changes each run: if you like the layout for this dataset, you can save the layout and load it next time:
#write.table(GSEA_abs_layout, paste0(input.path,"circlepack_layout.tsv"), sep="\t", quote = F, row.names = F)

main_plot =
    ggraph(GSEA_abs_layout %>% arrange(Type))+
    geom_node_circle(aes(fill=Type, color=Type, group=leaf, linetype=as.factor(Type)))+
    theme_void()+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type])+
    scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type])+
    coord_equal(clip='off')+
    theme(legend.position = "none")+
    ggraph::geom_node_label(aes(label=str_wrap(label,23)),alpha=0.6,lineheight = 0.7, size=3)

TP_FN_legend = layout[[i]] %>% filter(Type %in% c("TP", "FN", "Unable to produce biomass when KO'd")) %>%
    mutate(Type = fct_relevel(Type, c("TP", "FN", "Unable to produce biomass when KO'd"))) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type, linetype=as.factor(Type)))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white", name="Predicted",
                      labels=c("True Positive", "False Negative", 
                               "Infeasible"))+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Predicted",
                      labels=c("True Positive", "False Negative", 
                               "Infeasible"))+
  scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type], 
                        labels=c("True Positive", "False Negative", 
                               "Infeasible"), name="Predicted")+
    theme(legend.margin = margin(c(-0.5,0,0,0)))
    
cat_legend = layout[[i]] %>% filter(!Type %in% c("TP", "FN", "Unable to produce biomass when KO'd")) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white", name="Pathway category")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Pathway category")+
    theme(legend.margin = margin(c(-0.5,0,0,0)))
  
p3 = plot_grid(
    main_plot,
    plot_grid(get_legend(TP_FN_legend),
    get_legend(cat_legend),nrow = 2, align = "hv", rel_heights = c(0.3,0.7))
    ,nrow = 1, rel_widths = c(0.6,0.4))
p3
GSEA_abs_no_leg = main_plot
```
## ORA

```{r}
i=methods.indexes["ORA"]
ORA_layout = layout[[i]] %>% select(-x,-y, -r) %>% left_join(layout[[base_layout]] %>% select(x, y,r, name), by="name")

main_plot =
    ggraph(ORA_layout%>% arrange(Type))+
    geom_node_circle(aes(fill=Type, color=Type, group=leaf, linetype=as.factor(Type)))+
    theme_void()+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type])+
    scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type])+
    coord_equal(clip='off')+
    theme(legend.position = "none")+
    ggraph::geom_node_label(aes(label=str_wrap(label,23)),alpha=0.6,lineheight = 0.7, size=3)

TP_FN_legend =
    layout[[i]] %>% filter(Type %in% c("TP", "FN", "Insufficient significant metabolites", "Unable to produce biomass when KO'd", "TP_WT_default", "FN_WT_default")) %>% mutate(Type = fct_relevel(Type, c("TP", "FN", "Insufficient significant metabolites", "Unable to produce biomass when KO'd"))) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type, linetype=as.factor(Type)))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "white", name="Predicted",
                      labels=str_wrap(c("True Positive", "False Negative", 
                               "Insufficient significant metabolites",
                               "Infeasible"
                               ), 23))+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Predicted",
                      labels=str_wrap(c("True Positive", "False Negative",  
                               "Insufficient significant metabolites",
                               "Infeasible"), 23))+
  scale_linetype_manual(values=linetype_dict[pathway_counts_run[[i]]$Type], 
                        labels=str_wrap(c("True Positive", "False Negative", 
                               "Insufficient significant metabolites",
                               "Infeasible"
                               ), 23), name="Predicted")

cat_legend = layout[[i]] %>% filter(!Type %in% c("TP", "FN", "Insufficient significant metabolites", "Unable to produce biomass when KO'd")) %>% ggraph()+
    geom_node_circle(aes(fill=Type, group=depth, color=Type))+
    scale_fill_manual(values = pathway_cat_cols_dict_light[pathway_counts_run[[i]]$Type],na.value = "lightgrey", name="Pathway category")+
    scale_color_manual(values = pathway_cat_cols_line_dict[pathway_counts_run[[i]]$Type], name="Pathway category")#+
    # theme(legend.margin = margin(c(0,0,0,0)))
  
p1 = plot_grid(
    main_plot,
    plot_grid(get_legend(TP_FN_legend),
    get_legend(cat_legend),nrow = 2, align = "hv", rel_heights = c(0.3,0.7))
    ,nrow = 1, rel_widths = c(0.6,0.4))
p1
ORA_no_leg = main_plot
```
```{r}
ggsave("~/Downloads/ORA_pink.png", width = 8, height = 6, dpi=300, bg="white")
```

## Fig2
### Legend
Made manually, can need adjusting visually
```{r}
if (model_name == "Recon2.2"){
  circle_div = 18
}else{
  circle_div = 25
}

max_size = max(layout[[1]]$n)
min_size = min(layout[[1]]$n)
scale_legend =
  layout[[1]] %>% 
  ggplot()+
  geom_point(aes(size=n), x=1, y=1, pch=21, fill="white", color="#3b3b3b")+
  scale_size_continuous(breaks =c(min_size,100,max_size), name="Pathway size", range=c(3,max_size/circle_div))+
  theme_minimal()
  # theme(legend.direction = "horizontal", legend.title.position = "top")

circle_legend =
  plot_grid(plot_grid(
    get_legend(TP_FN_legend),
    NULL,
    get_legend(scale_legend), 
    nrow=3, align="hv", rel_heights = c(0.48,0.005,0.48)),
    
    get_legend(cat_legend),
    
    nrow = 1, align = "v")+theme(plot.margin = margin(0,2,0,0, "cm"))
#cat_legend+theme(plot.margin = margin(0.5,1,0.5,0.5, "cm"))
```

```{r}
plot_grid(ORA_no_leg, circle_legend, NULL,NULL, GSEA_abs_no_leg, GSEA_raw_no_leg,  labels = c("A","","","","B","C"), nrow=3, rel_heights = c(0.45,0.03,0.45),
          hjust=c(-0.5,0,0,0,-0.3,-0.3))

ggsave(paste0("../../plots/",model_name,"/Fig2.png"), bg="white", width=11, height = 10, dpi=300)
#ggsave("../../plots/Fig2-2-no-bg.png", width=11, height = 10, dpi=300)
```


# Fig3
## Barplot
```{r}
all_methods_TP_FN_long = lapply(c(1:3), function(i){
  all_methods_TP_FN_cat[[i]] %>% mutate(method = method_names[i])
})
all_methods_df = do.call(rbind, all_methods_TP_FN_long)

all_methods_df_FPR_TNR = all_methods_df %>% pivot_longer(c(FPR, TNR)) %>% select(-Type) %>% group_by(method, name) %>% 
  summarise(value = mean(value, na.rm=T)) %>% mutate(group=2) %>% rename("Type" = name) %>% 
  ungroup() %>% 
  mutate(method = fct_relevel(method,method_names))

TP_FN_plot =
  all_methods_df %>% 
  filter(!Type %in% c("Unable to produce biomass when KO'd", "Unable to produce biomass in WT")) %>% 
  group_by(method, Type) %>% 
  summarise(n = n()) %>% 
  mutate(total = sum(n), value = n/total) %>% 
  ungroup() %>% 
  mutate(Type = fct_relevel(Type, c("Insufficient significant metabolites","FN","TP")),
         method = fct_relevel(method, paste0(model_name, "_", 
                                             c("ORA_pval_005_metrics_external", 
                                               "GSEA_absolute_metric_external", 
                                               "GSEA_raw_metric_external"), 
                                             "_with_WT_default_all")), group=1) %>% 
  select(method, Type, value, group) %>% 
  ggplot(aes(x=method, y= value, fill=Type))+
  geom_bar(stat="identity", col="white")+
  geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5), col="white", size=4)+
  theme_minimal_hgrid()+
  scale_fill_manual(values = pathway_cat_cols_dict[all_methods_df$Type], 
                    labels=c(str_wrap("Insufficient significant metabolites",50),"FN","TP"))+
  scale_x_discrete(labels = c("ORA", "GSEA\nunsigned", "GSEA\nsigned"))+
  xlab("")+
  ylab("Proportion")+
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        legend.position = "right", legend.justification = "center",
        legend.direction = "vertical", legend.title = element_blank())

FPR_TNR_colours = c("#ed978c", "#8a82ed")
names(FPR_TNR_colours) = c("FPR", "TNR")

FPR_TNR_plot = all_methods_df_FPR_TNR %>% 
ggplot(aes(x=method, y= value, fill=Type))+
  geom_bar(stat="identity", col="white")+
  geom_text(aes(label=round(value,2)), position = position_stack(vjust=0.5), col="white", size=4)+
  theme_minimal_hgrid()+
  scale_fill_manual(values = FPR_TNR_colours[all_methods_df_FPR_TNR$Type], labels=c("FP", "TN"))+
  scale_x_discrete(labels = c("ORA", "GSEA\nunsigned", "GSEA\nsigned"))+
  xlab("")+
  ylab("Counts")+
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(),
        legend.position = "right", legend.justification = "center",
        legend.direction = "vertical", legend.title = element_blank())

ggarrange(TP_FN_plot, FPR_TNR_plot, widths = c(0.55,0.45), labels = "AUTO", align="h")
```

## Z-scores
```{r}
z.scores.long = z.scores %>% pivot_longer(-Metab) %>% 
  left_join(pathway_dict, by = c("name" = "Group")) %>% select(-name) %>% 
  left_join(metabolites, by = c("Metab" = "ID"))
length.to.strip = str_length(exch.metab.pattern)+1
pathways_no_comp2 = pathways %>% mutate(metabolite = stringr::str_sub(metabolite, end=-length.to.strip)) %>% unique() %>% 
  rename("Pathway" = subsystem)

zthr=50
zscores.kod.pathways = pathways_no_comp2 %>% 
  left_join(z.scores.long, by = c("Pathway", "metabolite" = "metabID")) %>% 
  filter(!is.na(value)) %>% 
  # group_by(Pathway) %>% 
  # summarise(value = mean(value)) %>%
  left_join(all_methods_TP_FN_long_df %>% 
              mutate(Method = str_replace_all(Method, method_names_clean))%>% 
              filter(Method == "GSEA abs"), 
            by = c("Pathway")) %>% 
  group_by(Pathway) %>% 
  reframe(sd =sd(value), metabolite, Pathway, value, Type, n=n()) %>% 
  filter(!is.na(Type)) %>% 
  arrange(sd) %>% 
  mutate(Pathway = fct_inorder(Pathway)) %>% 
  mutate(value = ifelse(value >= zthr, zthr, 
                        ifelse(value <= -zthr, -zthr, value)))


zscores.kod.pathways %>% 
  filter(value >= zthr | value <= -zthr) 
# 6 metabolites modified for plotting purposes

zscores.plot = zscores.kod.pathways %>%
  ggplot(aes(y=(value), x= Pathway))+
  # geom_bar(aes(x=Pathway, y=n, fill=Type), stat="unique", alpha=0.4)+
  geom_boxplot(aes(col=Type),outliers=T,outlier.alpha = 0.2)+
  scale_colour_manual(values = pathway_cat_cols_dict[zscores.kod.pathways$Type], name = "GSEA unsigned\nprediction")+
  theme_minimal()+
  theme(axis.text.x = element_blank())+
  xlab("KO'd pathways (sorted by std)")+
  ylab("KO'd pathway's\nmetabolite z-scores")+
  guides(col=guide_legend(reverse=TRUE))
zscores.plot
```

## Fig3
```{r}
plot_grid(plot_grid(TP_FN_plot, NULL,FPR_TNR_plot, rel_widths = c(0.52,0.05,0.48), 
                    ncol = 3, labels=c("A","B")),
  nrow=2,
  zscores.plot,
  labels = c("", "C"), rel_heights = c(0.33,0.33, 0.33))
ggsave(paste0("../../plots/",model_name,"/Fig3.png"), dpi=300, bg="white", width=10, height=7)
```

```{r}
TP_FN_plot
ggsave(paste0("../../plots/",model_name,"/Fig3_vposter.png"), dpi=300, bg="white", width=6.5, height=4)
```


# Fig4
## Exchangeable metabolites ratio
Exchangeable metabolites ratio is defined by:
n. of unique exchangeable metabolites (not counting side compounds)
/
n. of total unique metabolites in the pathways (not counting side compounds)

This penalises side compounds as even if they are exchangeable, they do not add information to the metabolic profile. 
(most of the time)
```{r}
length.to.strip = str_length(exch.metab.pattern)+1

exchange_metabolites = pathways %>% filter(endsWith(metabolite, exch.metab.pattern)) %>% 
  mutate(metabolite = str_sub(metabolite, end = -length.to.strip)) %>% 
  filter(!(metabolite %in% (side_compounds %>%
           mutate(Identifier = str_remove(Identifier, pattern="^M_")) %>%
           pull(Identifier)))) %>%
  pull(metabolite) %>% unique()
  

exchangeable_metabolites_per_pathway = pathways %>% 
  mutate(metabolite = str_sub(metabolite, end = -length.to.strip)) %>% group_by(subsystem) %>% 
  filter(metabolite %in% exchange_metabolites) %>% unique() %>% summarise(Exchangeable = length(metabolite)) %>% 
  left_join(pathways %>% 
              mutate(metabolite = str_sub(metabolite, end = -length.to.strip)) %>% 
              filter(!(metabolite %in% (side_compounds %>%
                                          mutate(Identifier = str_remove(Identifier, pattern="^M_")) %>%
                                          pull(Identifier)))) %>% 
              unique() %>%  
              group_by(subsystem) %>% 
              summarise(Total=length(metabolite)), by = "subsystem") %>% 
  mutate(ExchangeableRatio = Exchangeable/Total)

if (exists("pathways.to.exclude")){
  exchangeable_metabolites_per_pathway = exchangeable_metabolites_per_pathway %>%
  filter(!subsystem %in% pathways.to.exclude)
}

for (i in 1:length(all_methods_TP_FN)){
  exchangeable_metabolites_per_pathway = exchangeable_metabolites_per_pathway %>% 
    left_join(all_methods_TP_FN[[i]] %>% select(Subsystem.name, Type) %>% rename(!!filenames[[i]] := "Type"), by = c("subsystem" = "Subsystem.name"))
}

exchangeable_m_filtered = exchangeable_metabolites_per_pathway %>% 
  left_join(pathway_categories %>% select(Pathway, ManualCategoryLvl1), by = c("subsystem" = "Pathway")) %>% 
  filter(!subsystem %in% misc_blocked_pathways)

exchangeable_m_filtered$subsystem = factor(exchangeable_m_filtered$subsystem, levels=exchangeable_m_filtered %>% arrange(ExchangeableRatio) %>% pull(subsystem))

exchangeable_m_filtered = exchangeable_m_filtered %>% 
  pivot_longer(!!method_names) 
exchangeable_m_filtered$name = factor(exchangeable_m_filtered$name, levels=method_names)

exchangeable_m_filtered = exchangeable_m_filtered %>% 
  filter(!subsystem %in% infeasible_pathways_ko) %>% 
  mutate(value = ifelse(startsWith(as.character(name), "ORA") & is.na(value),"Insufficient significant metabolites", value)) %>% 
  filter(Total >= 3)

# Exchangeable Ratio bar plot
ex_ratio_plot =
  exchangeable_m_filtered %>% 
  filter(endsWith(as.character(name), "all")) %>% 
  select(subsystem, ExchangeableRatio) %>% unique() %>% 
  ggplot()+
  geom_bar(aes(x=ExchangeableRatio, y = subsystem), stat = "identity", fill = basic_colour)+
  theme_minimal()+
    theme(axis.text.y = element_text(size=7))+
  ylab("")+
  xlab("Exchangeable metabolites ratio")


exchangeable_m_filtered_clean = exchangeable_m_filtered %>% 
  filter(endsWith(as.character(name), "all"))%>% 
    mutate(name=str_replace_all(name, method_names_clean))
exchangeable_m_filtered_clean$name = factor(exchangeable_m_filtered_clean$name, levels = method_names_clean)

method_names_clean2 = c("ORA", "GSEA\nunsigned", "GSEA\nsigned")
names(method_names_clean2) = method_names_clean

# TP and FN heatmap
ex_TP_FN_plot =
  exchangeable_m_filtered_clean %>% 
    mutate(name = fct_relevel(name, c("GSEA abs","GSEA raw", "ORA"))) %>%
  filter(name == "GSEA abs") %>% 
    ggplot()+
  facet_wrap(~name,nrow = 1, scales="free_x", labeller = as_labeller(method_names_clean2))+
      geom_tile(aes(x=name, y= subsystem, fill = value), colour = "white", linewidth = 0.7)+
      scale_fill_manual(values = pathway_cat_cols_dict[exchangeable_m_filtered_clean$value], na.value = "lightgrey")+
      # scale_x_discrete(position="top", label= c("GSEA unsigned", "GSEA signed","ORA"))+
      theme_minimal()+
      theme(axis.text.y = element_blank(),
            axis.title = element_blank(), 
            axis.ticks = element_blank(),
            legend.position = "none", 
            strip.text = element_text(angle=45, hjust = 0.5, vjust = 0.4),
            axis.text.x=element_blank(),
            # axis.text.x = element_text(angle=45, hjust = 0.1, vjust = 0),
            panel.grid = element_blank())

# TP and FN legend
ex_TP_FN_plot_legend =
  exchangeable_m_filtered %>% 
  filter(name=="Human1_GSEA_absolute_metric_external_with_WT_default_all") %>% 
  mutate(value = fct_relevel(value, c("TP", "FN"))) %>%
  ggplot()+
  geom_tile(aes(x=name, y= subsystem, fill = value))+
  scale_fill_manual(values = pathway_cat_cols_dict[exchangeable_m_filtered_clean$value], name="Predicted", na.value = "lightgrey",
                    labels=c("True Positive", "False Negative"))+
  theme_void()+
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank())

# Pathway category plot
ex_category_plot =
  exchangeable_m_filtered %>% 
  select(subsystem, ManualCategoryLvl1) %>% 
    mutate("label" = "Pathway\ncategory") %>% 
  ggplot()+
  geom_tile(aes(x = label, y= subsystem, fill = ManualCategoryLvl1), linewidth = 0.7, colour = "white")+
  scale_fill_manual(values = pathway_cat_cols_dict[exchangeable_m_filtered$ManualCategoryLvl1])+
    scale_x_discrete(position="top")+
  theme_minimal()+
  theme(axis.text.y = element_blank(), axis.title = element_blank(), axis.ticks = element_blank(),
        legend.position = "none",
        axis.text.x = element_text(angle=45, hjust = 0.1, vjust = 0), panel.grid = element_blank())

# Pathway category plot legend
ex_category_plot_legend =
  exchangeable_m_filtered %>% 
  select(subsystem, ManualCategoryLvl1) %>% 
  ggplot()+
  geom_tile(aes(x=1, y= subsystem, fill = ManualCategoryLvl1))+
  scale_fill_manual(values = pathway_cat_cols_dict[exchangeable_m_filtered$ManualCategoryLvl1], name = "Pathway category")+
  theme_void()+
  theme(axis.text = element_blank(), axis.title = element_blank(), axis.ticks = element_blank())+
  guides(fill = guide_legend(override.aes = list(alpha = 1))) 

# Join all the plots together
exch_plot = plot_grid(
  plot_grid(
    ex_ratio_plot,
    ex_TP_FN_plot,
    ex_category_plot,
    rel_widths = c(0.83,0.07,0.07), nrow = 1, align = 'h', axis = "tb"),
  plot_grid(NULL,
    get_legend(ex_TP_FN_plot_legend),
    get_legend(ex_category_plot_legend),
    NULL,
      rel_heights = c(0.25,0.25,0.25,0.25),
      align = 'hv',
      nrow = 4, ncol = 1),
rel_widths = c(0.8, 0.2), nrow = 1)
exch_plot
```
## Fig4
```{r}
plot_grid(exch_plot)#, 
          # labels = c("","", "C"), 
          # rel_widths = c(0.2,0.05,0.8))
ggsave(paste0("../../plots/",model_name,"/Fig4.png"), dpi=300, bg="white", width=15, height=12)
```


# Fig5

## Uniqueness
```{r}
thr=2
uniqueness = z.scores %>% pivot_longer(-Metab) %>% 
  left_join(pathway_dict, by = c("name" = "Group")) %>% select(-name) %>% 
  rename("zscores" = value) %>% 
  group_by(Metab) %>% 
  # reframe(zscores = (value - mean(value)) / sd(value), Pathway) %>% 
  filter(zscores > thr | zscores < -thr) %>% 
  left_join(metabolites %>% select(-metabID), by = c("Metab" = "ID")) %>% 
  select("Pathway","Name","zscores") %>% #pull(Pathway) %>% table() %>% sort(decreasing = T)
  # mutate(direction = ifelse(zscores > 0, "pos", "neg")) %>% 
  group_by(Name) %>% 
  summarise(n=n(), Pathway) %>% #pull(n) %>% table() %>% sort(decreasing = T)
  group_by(Pathway) %>% 
  mutate(score = sum(1/n), n.metab = sum(n()), unique = sum(n==1), unique_score = unique/n.metab) %>% 
  select(-Name, -n) %>% 
  unique() %>% 
  arrange(desc(score))
```

## Adding significance
```{r}
z.scores.long = z.scores %>% pivot_longer(-Metab) %>% 
  left_join(pathway_dict, by = c("name" = "Group")) %>% select(-name) %>% 
  left_join(metabolites, by = c("Metab" = "ID"))

pathways_no_comp2 = pathways %>% mutate(metabolite = stringr::str_sub(metabolite, end=-length.to.strip)) %>% unique() %>% 
  rename("Pathway" = subsystem)

mean.zscores = 
  all_methods_TP_FN_long_df %>% 
  left_join(pathways_no_comp2, by = c("Pathway")) %>% 
  left_join(z.scores.long , by = c("Pathway", "metabolite" = "metabID")) %>% 
  filter(!is.na(value)) %>% group_by(Pathway, Method) %>% 
  summarise(mean = mean(abs(value)), Type) %>% unique() 

non.uniques = data.frame("Pathway" = exchangeable_metabolites_per_pathway$subsystem[!exchangeable_metabolites_per_pathway$subsystem %in% uniqueness$Pathway],
           "score"= 0, "n.metab" = 0, "unique" = 0, "unique_score" = 0)


variables.all = c("PathwaySize", "ExchangeableRatio", "ProfileUniquenessScore", "ClusteringCoefficient", "Degree", "NeighborhoodConnectivity")

boxplot.df = exchangeable_metabolites_per_pathway %>% filter(!(subsystem %in% misc_pathways)) %>% 
  left_join(uniqueness %>% rbind(non.uniques), by = c("subsystem"= "Pathway")) %>% 
  select(-paste0(model_name,"_GSEA_absolute_metric_external_with_WT_default_all"), -paste0(model_name,"_ORA_pval_005_metrics_external_with_WT_default_all"), -paste0(model_name,"_GSEA_raw_metric_external_with_WT_default_all")) %>% 
  left_join(mean.zscores, by = c("subsystem"= "Pathway")) %>% 
  rename("ProfileUniquenessScore" = score, "PathwaySize" = Total) %>% 
  select(subsystem, ProfileUniquenessScore, ExchangeableRatio, PathwaySize, Method, Type) %>% 
  left_join(network_stats %>% select(name, AverageShortestPathLength, BetweennessCentrality, ClosenessCentrality, ClusteringCoefficient, Degree, NeighborhoodConnectivity, Stress, TopologicalCoefficient), by = c("subsystem" = "name")) %>% 
  mutate(ProfileUniquenessScore = ifelse(is.na(ProfileUniquenessScore), 0, ProfileUniquenessScore)) %>% 
  filter(Degree != 0) %>%
    select(-c(AverageShortestPathLength ,  Stress, ClosenessCentrality,TopologicalCoefficient,BetweennessCentrality))%>%
  filter(PathwaySize > 5) %>%
  mutate(Method = str_replace_all(Method, method_names_clean)) %>% 
  filter(!is.na(Method)) %>%
  pivot_longer(-c(subsystem, Method, Type)) %>% 
  mutate(name = fct_relevel(name, variables.all),
         Method = fct_relevel(Method,method_names_clean))

# Build the boxplot to get boxplot auto outlier data
boxplot.data.all = ggplot_build(boxplot.df %>% 
  ggplot(aes(x=Type, y=value))+
  geom_boxplot(aes(col=Type, fill=Type), outliers = F, alpha=0.4)+
  facet_grid(Method~name, scales = "free",switch = "y"))$data[[1]]

# Get the max value of each facet
maxes.all = boxplot.df %>% 
  group_by(name, Method) %>% 
  summarise(y_position=max(value)) %>% 
  cbind(PANEL = c(1,7,13,2,8,14,3,9,15,4,10,16,5,11,17,6,12,18)) %>% # Set the facets manually...
  left_join(boxplot.data.all %>% 
              select(ymax, PANEL) %>% 
              group_by(PANEL) %>% 
              filter(ymax == max(ymax)) %>% 
              mutate(PANEL = as.numeric(as.character(PANEL))) %>% unique(), by=c("PANEL")) %>% 
  rename("y.position" = "ymax")

# Test each TP vs FN statistically
anno_df.all = compare_means(value ~ Type, group.by = c("name", "Method"), data = boxplot.df, method = "wilcox.test", p.adjust.method = "fdr") %>%
 mutate(p.adj = format.pval(p.adj, digits = 2)) %>%
  filter(p.signif != "ns") %>% 
  left_join(maxes.all, by=c("name", "Method")) %>% 
  mutate(y.position = y.position *1.1)

label.clean = c(str_wrap(str_replace_all(variables.all,"(.)(?=[A-Z])", "\\1 "),50), method_names_clean2)
names(label.clean) = c(variables.all, names(method_names_clean2))

all.boxplots = boxplot.df %>% 
  ggplot(aes(x=Type, y=value))+
  geom_boxplot(aes(col=Type, fill=Type), outliers = F, alpha=0.4)+
  facet_grid(Method~name, scales = "free",switch = "y", labeller = as_labeller(label.clean))+
  theme_minimal()+
  theme(strip.text.y.left = element_text(angle = 0,size = 8),
        axis.text.y = element_blank(),
        panel.spacing.x = unit(1, "lines"),
        legend.position = "right", legend.title = element_blank())+
  scale_fill_manual(values = pathway_cat_cols_dict[boxplot.df$Type])+
  scale_colour_manual(values = pathway_cat_cols_dict[boxplot.df$Type])+
  xlab("")+
  ylab("")+
  stat_pvalue_manual(anno_df.all, label="p.signif", coord.flip = T,tip.length = 0)+
  guides(col=guide_legend(reverse=TRUE), fill=guide_legend(reverse=TRUE))+
  coord_flip(clip = "off")
all.boxplots
```

```{r}
boxplot.df %>% 
  filter(Method == "GSEA abs" & name=="ExchangeableRatio") %>% 
  ggplot(aes(x=Type, y=value))+
  geom_boxplot(aes(col=Type, fill=Type), outliers = F, alpha=0.4)+
  # facet_grid(Method~name, scales = "free",switch = "y", labeller = as_labeller(label.clean))+
  theme_minimal()+
  theme(#strip.text.y.left = element_text(angle = 0,size = 8),
        # axis.text.y = element_blank(),
        panel.spacing.x = unit(1, "lines"),
        legend.position = "right", legend.title = element_blank())+
  scale_fill_manual(values = pathway_cat_cols_dict[boxplot.df$Type])+
  scale_colour_manual(values = pathway_cat_cols_dict[boxplot.df$Type])+
  xlab("")+
  ylab("")+
  stat_pvalue_manual(anno_df.all %>% filter(Method == "GSEA abs" & name=="ExchangeableRatio"), label="p.signif", coord.flip = F,tip.length = 0)+
  guides(col=guide_legend(reverse=TRUE), fill=guide_legend(reverse=TRUE))+
  coord_cartesian(clip = "off")
```

```{r}
ggsave("~/Downloads/exch_ratio_boxplot_poster.png", width = 2.5, height =3, dpi=300, bg="white")
```


## PCA
```{r}
z.scores.long = z.scores %>% pivot_longer(-Metab) %>% 
  left_join(pathway_dict, by = c("name" = "Group")) %>% select(-name) %>% 
  left_join(metabolites, by = c("Metab" = "ID"))

pathways_no_comp2 = pathways %>% mutate(metabolite = stringr::str_sub(metabolite, end=-length.to.strip)) %>% unique() %>% 
  rename("Pathway" = subsystem)

mean.zscores = 
  all_methods_TP_FN_long_df %>% 
  filter(Method == paste0(model_name,"_GSEA_absolute_metric_external_with_WT_default_all")) %>% 
  select(Pathway, Type)  %>% 
  left_join(pathways_no_comp2, by = c("Pathway")) %>% 
  left_join(z.scores.long , by = c("Pathway", "metabolite" = "metabID")) %>% 
  filter(!is.na(value)) %>% group_by(Pathway) %>% 
  summarise(mean = mean(abs(value)), Type) %>% unique() 

non.uniques = data.frame("Pathway" = exchangeable_metabolites_per_pathway$subsystem[!exchangeable_metabolites_per_pathway$subsystem %in% uniqueness$Pathway],
           "score"= 0, "n.metab" = 0, "unique" = 0, "unique_score" = 0)


GSEA.NES = GSEA.abs.results %>% 
  # filter(Term == unique_id & NOM.p.val > 0.05) %>% 
  filter(Term == unique_id) %>% 
  select(Term, NES)

pca.data = exchangeable_metabolites_per_pathway %>% filter(!(subsystem %in% misc_pathways)) %>% 
  left_join(uniqueness %>% rbind(non.uniques), by = c("subsystem"= "Pathway")) %>% 
  left_join(mean.zscores, by = c("subsystem"= "Pathway")) %>% 
  rename("GSEA.abs" = paste0(model_name, "_GSEA_absolute_metric_external_with_WT_default_all"),
         "ProfileUniquenessScore" = score, "PathwaySize" = Total) %>% 
  select(subsystem, GSEA.abs, ProfileUniquenessScore, ExchangeableRatio, PathwaySize) %>% 
  left_join(network_stats %>% select(name, AverageShortestPathLength, BetweennessCentrality, ClosenessCentrality, ClusteringCoefficient, Degree, NeighborhoodConnectivity, Stress, TopologicalCoefficient), by = c("subsystem" = "name")) %>% 
  filter(!is.na(GSEA.abs))%>%
  # filter(!subsystem %in% c("Peptide metabolism", "Vitamin B2 metabolism", "Riboflavin metabolism"))%>%
  # select(-c(Stress,BetweennessCentrality,AverageShortestPathLength, Degree,NeighborhoodConnectivity, TopologicalCoefficient, score))# %>%

  mutate(ProfileUniquenessScore = ifelse(is.na(ProfileUniquenessScore), 0, ProfileUniquenessScore)) %>% 
  left_join(GSEA.NES, by = c("subsystem" = "Term")) %>% 
  mutate(NES = ifelse(GSEA.abs == "FN" & is.na(NES), -1.5, NES))%>% 
  filter(Degree != 0) %>%
    select(-c(AverageShortestPathLength ,  Stress, ClosenessCentrality,TopologicalCoefficient,BetweennessCentrality))%>%
  # Recon2.2:
    # select(-c(Stress, Degree,score, AverageShortestPathLength, Total, TopologicalCoefficient, NeighborhoodConnectivity))
  filter(PathwaySize > 5) %>%
  mutate(NES = ifelse(NES <0, min(abs(NES)), NES)) 


indiv = pca.data %>% select(-GSEA.abs, -NES) %>% 
  tibble::column_to_rownames("subsystem")

groups = pca.data %>% column_to_rownames("subsystem") %>% select(GSEA.abs, NES)

res.PCA = PCA(indiv,scale.unit = T) 

arrow_coord = fviz_pca_biplot(res.PCA)$layers[[6]]$data
# Radial shift function
rshift = function(r, theta, a=0.1, b=0.7) {
  r + a + b*abs(cos(theta))
}

# Calculate shift
arrow_coord = arrow_coord %>% 
  mutate(r = sqrt(x^2 + y^2),
         theta = atan2(y,x),
         rnew = rshift(r, theta),
         xnew = rnew*cos(theta),
         ynew = rnew*sin(theta))

pca.coord = res.PCA$ind$coord %>% as.data.frame() %>% 
  rownames_to_column("subsystem") %>% 
  left_join(pathway_categories %>% select(Pathway, ManualCategoryLvl1), by = c("subsystem" = "Pathway"))

pca_biplot =
  pca.coord %>% 
  ggplot(aes(x=Dim.1, y=Dim.2))+
  geom_segment(data = arrow_coord, aes(xend=x,yend=y, x=xstart,y=ystart), inherit.aes = F, arrow = arrow(length = unit(0.2, "cm")))+
  geom_point(aes(shape=factor(groups$GSEA.abs), 
                 fill=factor(groups$GSEA.abs)), size = 3, alpha=0.8)+
  scale_shape_manual(values = c(23,21), name = "GSEA prediction")+
  scale_fill_manual(values = pathway_cat_cols_dict[groups$GSEA.abs], name = "GSEA prediction")+
  scale_colour_manual(values = pathway_cat_cols_dict[groups$GSEA.abs], name = "GSEA prediction")+
  stat_ellipse(aes(col=factor(groups$GSEA.abs)))+
  coord_cartesian(clip="off")+
  theme_minimal()+
  theme(legend.position = "bottom",
          legend.title =element_text(margin=margin(0,20,0,0)),
          axis.title.y = element_text(hjust = 0.4),
        plot.margin = margin(0.4,0.2,0.2,0.7, "cm"))+
  geom_hline(yintercept = 0, linetype="dashed")+
  geom_vline(xintercept = 0, linetype="dashed")+
  xlab(paste0("Dim1 ", "(",  round(res.PCA$eig["comp 1", "percentage of variance"],1),"%)"))+
  ylab(paste0("Dim2 ", "(",  round(res.PCA$eig["comp 2", "percentage of variance"],1),"%)"))+
  geom_label(data = arrow_coord, 
             aes(x=xnew,y=ynew, label = str_wrap(str_replace_all(name,"(.)(?=[A-Z])", "\\1 "),50)), 
             inherit.aes = F,
             fill=alpha(c("white"), 0.5))
pca_biplot
```


## Fig5 
```{r}
pathway_graph = ggdraw() + draw_image(paste0("../../plots/",model_name,"/",model_name,"_pathway_graph_edited2.png"))
fig5 = plot_grid(all.boxplots,labels = c("A"),
                 plot_grid(pathway_graph, pca_biplot, ncol=2, rel_widths = c(0.45,0.55), labels = c("B","C")), 
                 nrow=2, rel_heights = c(0.35,0.65))
ggsave(plot=fig5, paste0("../../plots/",model_name,"/Fig5.png"), dpi=300, bg="white", width = 15, height=10)
```